Reducing Acceptance Marks in Emerson-Lei Automata by QBF-solving

### 2022-02-27 paper/main.tex
some progress in section 3

\newcommand{\Nset}{{\mathbb{N}_0}}
\newcommand{\todo}[1]{\textcolor{red}{#1}}
%\def\Inf{\ensuremath{\mathsf{Inf}}}
%\def\Fin{\ensuremath{\mathsf{Fin}}}
\DeclareMathOperator{\Inf}{\mathsf{Inf}}
\DeclareMathOperator{\Fin}{\mathsf{Fin}}
\newcommand{\minf}{\mathit{inf}}
\newcommand{\mks}{\mathit{mks}}
%\newcommand{\oM}{\widehat{M}}
\newcommand{\oM}{N}
\newcommand{\rem}{\mathit{nm}}
\newcommand{\mcycle}{\mathit{cycle}}
\newcommand{\msat}{\mathit{satisfies}}
\def\false{\mathit{false}}
\def\true{\mathit{true}}
\newcommand{\mA}{\mathcal{A}}
\newcommand{\mB}{\mathcal{B}}

@@

It remains to define the subformula $\mcycle(\vec{e})$ with the intended meaning
\[
  \vec{e}\textrm{ represents a cycle} \iff \mcycle(\vec{e}).
\]
In fact, only the direction ``$\Longrightarrow$'' is required for
correctness of our reduction method. If $\mcycle(\vec{e})$ holds for
some valuation of $\vec{e}$ that does not represent a cycle, then the
required equivalence
$\msat_\varphi(\vec{e}) \iff
\msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})$ on this valuation of
$\vec{e}$ adds a superfluous constraint on $\psi$ represented by
$\vec{i},\vec{f}$ and $\rem$ represented by $\vec{n}$. These
superfluous constraints can lead to loss of reduction opportunities,
but not to incorrectness. This observation allows us to trade the
precision of $\mcycle(\vec{e})$ for simplicity of the formula.

In the following, we define three versions of $\mcycle(\vec{e})$:
\begin{itemize}
\item $\mcycle_1(\vec{e})$ is a lightweight version, which puts
  only minimal requirements on $\vec{e}$. It does not use the
  information about automaton structure, but it comes with an
  interesting simplification of the whole formula $\Phi_{C,K}$.
\item $\mcycle_2(\vec{e})$ is a balanced version. It uses the
  information about the automaton structure to identify many
  valuations of $\vec{e}$ that are not cycles. It is relatively
  simple, but it is still satisfied by some valuations of $\vec{e}$
  not corresponding to cycles.
\item $\mcycle_3(\vec{e})$ is a strict version satisfied only by
  the valuations of $\vec{e}$ representing cycles. Unfortunately,
  it uses additional quantified variables. 
\end{itemize}

\subsection{Lightweight version $\mcycle_1(\vec{e})$}
The lightweight version is based on the fact that every cycle is nonempty.
Indeed, the only restriction imposed by
\[
  \mcycle_1(\vec{e})=\bigvee_{t\in\delta}e_t
\]
is that $\vec{e}$ represents a nonempty set.

The fact that this version do

>>>>

\newcommand{\Nset}{{\mathbb{N}_0}}
\newcommand{\todo}[1]{\textcolor{red}{#1}}
%\def\Inf{\ensuremath{\mathsf{Inf}}}
%\def\Fin{\ensuremath{\mathsf{Fin}}}
\DeclareMathOperator{\Inf}{\mathsf{Inf}}
\DeclareMathOperator{\Fin}{\mathsf{Fin}}
\newcommand{\minf}{\mathit{inf}}
\newcommand{\mks}{\mathit{mks}}
%\newcommand{\oM}{\widehat{M}}
\newcommand{\oM}{N}
\newcommand{\rem}{\mathit{nm}}
\newcommand{\mcycle}{\mathit{cycle}}
\newcommand{\msat}{\mathit{satisfies}}
\newcommand{\Te}{T_{\vec{e}}}
\def\false{\mathit{false}}
\def\true{\mathit{true}}
\newcommand{\mA}{\mathcal{A}}
\newcommand{\mB}{\mathcal{B}}

@@

It remains to define the subformula $\mcycle(\vec{e})$. Let $\Te$
denote the set of transitions represented by $\vec{e}$. The original
intended meaning of $\mcycle(\vec{e})$ is
\[
  \mcycle(\vec{e}) \iff \Te\textrm{ is a cycle}.
\]
In fact, only the direction ``$\Longleftarrow$'' is required for the
correctness of our reduction method. If $\Te$ is not a cycle and
$\mcycle(\vec{e})$ holds, then the required equivalence
$\msat_\varphi(\vec{e}) \iff
\msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})$ on this valuation of
$\vec{e}$ adds a superfluous constraint on $\psi$ represented by
$\vec{i},\vec{f}$ and $\rem$ represented by $\vec{n}$. These
superfluous constraints can lead to loss of reduction opportunities,
but not to incorrectness. This observation allows us to trade the
precision of $\mcycle(\vec{e})$ for simplicity of the formula.

In the following, we define three versions of $\mcycle(\vec{e})$:
\begin{itemize}
\item $\mcycle_1(\vec{e})$ is a lightweight version, which puts
  only minimal requirements on $\Te$. It does not use the
  information about automaton structure, but it comes with an
  interesting simplification of the whole formula $\Phi_{C,K}$.
\item $\mcycle_2(\vec{e})$ is a balanced version. It says that every
  transition in $\Te$ has to have a preceding and a succeeding
  transition in $\Te$, which is a necessary condition on cycles, but
  not a sufficient one.
\item $\mcycle_3(\vec{e})$ is a strict version saying that $\Te$ is a
  cycle. Unfortunately, it uses additional universally quantified
  variables corresponding to automata states. Transformation to prenex
  normal form turns the quantifiers to existential ones and the
  resulting formula thus contains quantifier alternation.
\end{itemize}

\subsection{Lightweight version $\mcycle_1(\vec{e})$}
The lightweight version is based on the fact that every cycle is nonempty.
Indeed, the only restriction imposed by
\[
  \mcycle_1(\vec{e})=\bigvee_{t\in\delta}e_t
\]
is that $\Te$ is nonempty.

With this lightweight version of $\mcycle(\vec{e})$, the formula
$\Phi_{C,K}$ actually says that
for every nonempty set
$T\subseteq\delta$, $T$ satisfies $\varphi$ if and only if $\rem(T)$ satisfies
$\psi$. Note that the formula is not affected by the placement of
transitions in the automaton. The only aspect of a transition $t$
reflected by the formula is its set of marks $\mks(t)$. Hence, we do
not have to distinguish between transitions with the same sets of
marks.

Formally, we define an equivalence on transitions such that for each
$t_1,t_2\in\delta$ we write $t_1\sim t_2$ whenever
$\mks(t_1)=\mks(t_2)$.
%This relation is clearly an equivalence.
We prove the following statement. 
\begin{lemma}
  Assume that there is a function $\rem$ and a formula $\psi$ such
  that for every nonempty set $T\subseteq\delta$ it holds that $T$
  satisfies $\varphi$ iff $\rem(T)$ satisfies $\psi$. Then we can
  safely assume that $\rem$ respects the quivalence $\sim$, that is
  $t_1\sim t_2\implies\rem(t_1)=\rem(t_2)$.
\end{lemma}
\begin{proof}
  \todo{Jan is here!}
\end{proof}

Formally, for each pair of transitions $t_1,t_2\in\delta$, we write
$t_1\sim t_2$ whenever $\mks(t_1)=\mks(t_2)$. This relation is clearly an equivalence.
Further, let us say that

We show that if this formula holds for some $\rem$ and $\psi$,
then it also holds for some $\rem'$ and $\psi$ such that $\rem'$
assigns the same sets of acceptance marks to transitions originally
marked by the same sets of marks.
